---
title: "Creative 3: Areas and Locations"
author: "Meghan Casey"
date: September 30, 2020
output: 
  html_document:
    theme: "cosmo"
    toc: true
    toc_depth: 3  
    toc_float: true    
    highlight: "tango"
---
### Load Libraries
```{r, message = FALSE}
library(sf)
library(tidyverse)
library(ggthemes)
library(units)
library(ggspatial)
library(nngeo)
```

### Load Data
##### I'm using data from Analyze Boston.

1. Boston Wards (polygon): wards
2. Public Libraries (point): lib
3. Main Street District (polygon): planning
4. Colleges + Universities (points): school

The Colleges and Universities dataset was downloaded from [http://bostonopendata-boston.opendata.arcgis.com/datasets/cbf14bb032ef4bd38e20429f71acb61a_2.kml}](http://bostonopendata-boston.opendata.arcgis.com/datasets/cbf14bb032ef4bd38e20429f71acb61a_2.kml}){target="_blank"}

```{r}
wards <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/398ee443f4ac49e9a0b7facf80afc20f_8.kml", quiet = TRUE)

lib <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/cb00f9248aa6404ab741071ca3806c0e_6.kml", quiet = TRUE)

planning <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/440c7ec0178d4c8593aecef7ea96bb4d_0.kml", quiet = TRUE)

school <- st_read("Colleges_and_Universities.kml", quiet = TRUE)
```

### Transforming Data
##### I'll be using the Massachusetts State Plane system
```{r}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

wards <- wards %>%
  st_transform(MA_state_plane)

lib <- lib %>%
  st_transform(MA_state_plane)

planning <- planning %>%
  st_transform(MA_state_plane)

school <- school %>%
  st_transform(MA_state_plane)
```

```{r}
ggplot(wards) +
  geom_sf(fill = "gray95", color = "gray40", size = 0.25) +
    geom_sf(data = planning, fill = "yellow3", color = NA, size = .60) +
  geom_sf(data = lib, color = "aquamarine4", size = 1.5) +
  geom_sf(data = school, color = "orangered3", size = 1.5, alpha = 0.5) +

  
  theme_pander()+
  annotation_scale()
```

### 1. Number and Proportion of Schools within a quarter mile of Public Libraries:
```{r}
lib_buffer <- st_buffer(lib, dist = 400) %>%
  st_union()

ggplot(lib_buffer) +
  geom_sf(fill = "aquamarine4", alpha = 0.25)+
  geom_sf(data = wards, fill = NA, alpha = 0.5) +
  theme_map()
```

```{r}
lib_school <- school[lib_buffer,]

ggplot(lib_buffer)+
  geom_sf(fill = "aquamarine4", alpha = 0.25)+
  geom_sf(data = wards, fill = NA, alpha = 0.5) +
  geom_sf(data = lib_school,
          color = "orangered3",
          size = 1.5)+
  theme_map()
```

```{r}
school <- school %>%
  st_join(lib_school) %>%
  mutate(by_lib = !is.na(Name.y))
```

```{r}
n_lib_school <- sum(school$by_lib)
n_lib_school
```

```{r}
n_school <- length(school$by_lib)

pct_lib_school <- n_lib_school / n_school

pct_lib_school
```
```{r}
left_side  <- st_bbox(school)$xmin
top_side <- st_bbox(school)$ymax

ggplot(wards) +
  geom_sf(fill = "white", color = "gray40", size = 0.25) +
  geom_sf(data = school, size = 1,
          aes(color = by_lib)) +
  scale_color_manual(values = c("aquamarine4", "orangered3"),
          name = "Boston Collages\nby Distance to\nPublic Library", 
          labels = c("No school withing 1/4 mile",
                     "School within 1/4 mile")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "bl",
                         style = north_arrow_fancy_orienteering()) +
  annotate(geom = "text", x = left_side, 
           y = top_side, 
           label = paste("Of the ", 
                         prettyNum(n_school, big.mark = ","),
                         " Colleges in Boston\n", 
                         prettyNum(n_lib_school, big.mark = ","),
                         " (", 
                         prettyNum(100*pct_lib_school, digits = 0),
                         "%) are within  a quarter\nmile of a public library.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = alpha("cadetblue4", 0.15)),
        legend.position = "left")
```

### 2. Number and Proportion of Colleges within Boston Wards:
```{r}
wards <- wards %>%
  mutate(n_school = lengths(st_covers(wards, school)))


ggplot(wards) +
  geom_sf(color = NA,
          aes(fill = n_school))+
  scale_fill_gradientn(colors = c("gray90", "salmon1", "salmon2", "salmon3", "salmon4"), name = "Boston Wards\nby Number\nof Colleges",
                      breaks = breaks <- seq(0, 20, by = 5)) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                       style = north_arrow_fancy_orienteering()) +
  
  theme_map() +
  theme(legend.position = "left")

```

### 3. Number and Proportion of Public Libraries within Boston Wards:
```{r}
wards <- wards %>%
  mutate(num_lib = lengths(st_covers(wards, lib)))


ggplot(wards) +
  geom_sf(color = NA,
          aes(fill = num_lib))+
  scale_fill_gradientn(colors = c("tomato3", "sandybrown", "khaki2", "palegreen3"), name = "Boston Wards\nby Number\nof Public Libraries",
                      breaks = breaks <- seq(0, 5, by = 1)) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                       style = north_arrow_fancy_orienteering()) +
  
  theme_map() +
  theme(legend.position = "left")
```

### 4. Number and Proportion of Libraries within Boston Main Street Districts:
```{r}
planning <- planning %>%
  mutate(num_lib = lengths(st_covers(planning, lib)))


ggplot(planning) +
  geom_sf(data = wards, fill = NA, alpha = 0.5) +
  geom_sf(color = NA,
          aes(fill = num_lib))+
  scale_fill_gradientn(colors = c("cyan4", "darkorange2"), name = "Boston Main Street Districts\nby Number\nof Public Libraries",
                      breaks = breaks <- seq(0, 2, by = 1)) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                       style = north_arrow_fancy_orienteering()) +
  
  theme_map() +
  theme(legend.position = "left")

```

### 5. Number and Proportion of Colleges within Boston Main Street Districts:
```{r}
planning <- planning %>%
  mutate(n_school = lengths(st_covers(planning, school)))

ggplot(planning) +
  geom_sf(data = wards, fill = NA, alpha = 0.5) +
  geom_sf(color = NA,
          aes(fill = n_school)) +
  scale_fill_gradientn(colors = c("tomato1", "yellow2", "olivedrab3"), name = "Boston Main Street Districts\nby number of Colleges",
                       breaks = breaks <- seq(0,10, by =1),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "colleges"))+
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                       style = north_arrow_fancy_orienteering()) + 
 theme_map() +
  theme(legend.position = "left")
```

### 6. Number and Proportion of Boston Main Street Districts that overlap with Boston Wards:
```{r}
wards <- wards %>%
  mutate(num_planning = lengths(st_overlaps(wards, planning))) %>%
  mutate(has_plan = num_planning > 0)
```
```{r}
n_plan_wards <- sum(wards$has_plan)
n_plan_wards
```
```{r}
left_side  <- st_bbox(planning)$xmin
top_side <- st_bbox(planning)$ymax

ggplot(planning) +
  geom_sf(fill = "gray40", color = NA) +
  geom_sf(data = wards,
          aes(fill = has_plan)) +
  scale_fill_manual(values = c("gray90", "yellow2"),
          name = "Boston Wards\nby presence of a Main Street District", 
          labels = c("Ward without an overlapping\nMain Street District",
                     "Ward with an overlapping\nMain Street District")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                       style = north_arrow_fancy_orienteering()) + 
  annotate(geom = "text", x = left_side, 
           y = top_side, 
           label = paste(n_plan_wards ,
                         "of Boston's", 
                         length(wards$Name),
                         "wards\noverlap with", 
                         "a main street district."),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(legend.position = "left")
```

### 7. Average Distance between Colleges and Public Libraries
```{r}
school <- school %>%
  mutate(lib_dist = st_nn(school, lib,
                             returnDist = TRUE)$dist) %>%
  mutate(lib_dist = as.numeric(lib_dist))
```

```{r}
avg_lib_dist <- mean(school$lib_dist)

avg_lib_dist
```

```{r}
right_side <- st_bbox(lib)$xmax
left_side  <- st_bbox(lib)$xmin
top_side <- st_bbox(lib)$ymax
bottom_side <- st_bbox(lib)$ymin


ggplot(wards) +
  geom_sf(fill = "white", color = "gray40") +
  geom_sf(data = school, size = 2,
          aes(color = lib_dist)) +
  scale_color_gradientn(colors = c("aquamarine3", "gold1", "salmon3"), name = 
                          "Boston Colleges\nby distance to Public Libraries") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "bl",
                       style = north_arrow_fancy_orienteering()) +
  annotate(geom = "text", x = left_side, 
           y = top_side - 100, 
           label = paste("On average, a Boston\nCollege is ", 
                         prettyNum(avg_lib_dist, digits = 3),
                         " meters\nfrom a public library.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(legend.position = "left")
```

### 8. Number and Proportion of Boston Wards that contain Colleges:
```{r}
wards <- wards %>%
  mutate(area = set_units(st_area(wards), km^2)) %>%
  mutate(school_dens = as.numeric(n_school / area))

ggplot(wards) +
  geom_sf(color = NA,
          aes(fill = school_dens)) +
  scale_fill_gradientn(colors = c("gray97", "cadetblue4", "darkslategrey"), name = "Boston Wards by\nCollege Density",
          breaks = breaks <- seq(0, 6, by = 1),
          labels = paste(prettyNum(breaks, big.mark = ","),
                                      "colleges per square km")) +
    annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                       style = north_arrow_fancy_orienteering()) + 
  theme_map() +
  theme(legend.position = "left")

```